Quantifying Heterogeneity in SAR Imagery with the Rényi Entropy

### 2025-06-03 Text/Articles/RENYI_ENTROPY_SAR_2025/Reviews/Responses-Reviewer1.tex

Additive noise is not considered in the model because it is usually very small in SAR images that have been properly calibrated.
 In our experiments, we use SAR images in intensity format (e.g., HH) that are radiometrically and geometrically corrected. 
Because the method analyzes small windows and uses non-parametric entropy estimation, it is relatively robust to small residual noise that may remain after standard preprocessing, such as calibration and terrain correction. 
We do not apply any additional filtering in order to keep the texture information that is important for detecting heterogeneity.

\end{responsebox}

\vspace{2em}
\begin{reviewbox}{Comment 5}
5. Fig 6, 7 and 8 show the detection of heterogeneous areas. However, a statistical measure indicating the detection performance is absent.
\end{reviewbox}

\begin{responsebox}{Response}
In the revised manuscript, we have included a quantitative performance evaluation to complement the visual comparison presented in Figs.~5a, 6a, and 7a (corresponding to Figs.~6a, 7a, and 8a in the original submission).

Specifically, we manually defined eight polygonal regions of interest (ROIs) in each SAR image: four clearly homogeneous areas ( water, flat croplands) and four clearly heterogeneous areas (urban centers). Each ROI was labeled as either homogeneous (class 0, shown in blue) or heterogeneous (class 1, in red).

@@

Additionally, to assess performance across all decision thresholds, we computed ROC curves and the area under the curve (AUC) for each method.

\begin{quote}
\textcolor{blue}{These additions appear in Section~IV-C of the revised manuscript.
}
\end{quote}
\end{responsebox}

\vspace{9em}
\begin{reviewbox}{Comment 6}
6. The color bars and the font size of the texts needs to be improved so that it can be clearly visible.
\end{reviewbox}
\begin{responsebox}{Response}

We improved Figures~6–8 (now Figs.~5–7)  by increasing the font size of labels, standardizing the size and placement of the color bars, and updating the colormap in Figure 8 to enhance visual contrast and improve the distinction between low and high $p$-values.

### 2025-06-03 Text/Articles/RENYI_ENTROPY_SAR_2025/Reviews/Responses-Associate-Editor.tex

\end{quote}

\end{responsebox}

\begin{reviewbox}{Comment 3}
3. Figures (Figs. 6–8) are informative but could benefit from quantitative evaluation. Add comparative metrics beyond visual comparison.
\end{reviewbox}

\begin{responsebox}{Response}
We agree with the reviewer and have added a quantitative performance evaluation to the revised manuscript. Specifically, we manually selected eight regions of interest (ROIs) in each SAR image: four clearly homogeneous areas (e.g., water, croplands) and four heterogeneous areas (e.g., urban centers). Each ROI was labeled as class~0 (blue) or class~1 (red), as illustrated in Figs.~5a, 6a, and 7a (corresponding to Figs.~6a, 7a, and 8a in the original submission).

@@

These ROIs were rasterized to create a partial ground-truth mask. Binary decision maps were obtained by thresholding the $p$-value maps at $ 0.05$ significance level, assigning 1 to pixels with $p < 0.05$ (heterogeneous) and 0 otherwise (homogeneous). These predictions were compared to the ground-truth labels, and standard metrics F1-score, kappa coefficient, and overall accuracy were computed. The results are reported in Table~III.

To further support the evaluation, we computed ROC curves and the area under the curve (AUC) for each method in Fig. 9.

\begin{quote}
\textcolor{blue}{These additions are included in Section~IV-C of the revised manuscript. }
\end{quote}

### 2025-06-03 Text/Articles/RENYI_ENTROPY_SAR_2025/Reviews/Responses-Reviewer1.tex

Additive noise is not considered in the model because it is usually very small in SAR images that have been properly calibrated.
 In our experiments, we use SAR images in intensity format (e.g., HH) that are radiometrically and geometrically corrected. 
Because the method analyzes small windows and uses non-parametric entropy estimation, it is relatively robust to small residual noise that may remain after standard preprocessing, such as calibration and terrain correction. 
We do not apply any additional filtering in order to keep the texture information that is important for detecting heterogeneity.

\end{responsebox}

\vspace{2em}
\begin{reviewbox}{Comment 5}
5. Fig 6, 7 and 8 show the detection of heterogeneous areas. However, a statistical measure indicating the detection performance is absent.
\end{reviewbox}

\begin{responsebox}{Response}
In the revised manuscript, we have included a quantitative performance evaluation to complement the visual comparison.

Specifically, we manually defined eight polygonal regions of interest (ROIs) in each SAR image: four clearly homogeneous areas ( water, flat croplands) and four clearly heterogeneous areas (urban centers). Each ROI was labeled as either homogeneous (class 0, shown in blue) or heterogeneous (class 1, in red),  as illustrated in Figs.~5a, 6a, and 7a (corresponding to Figs.~6a, 7a, and 8a in the original submission).

@@

These ROIs were then rasterized to create a partial ground-truth mask. 
We thresholded the $p$-value maps at $0.05$ significance level to generate binary decision maps: values below this threshold were labeled as heterogeneous (1), and values above or equal to the threshold as homogeneous (0). 
The resulting prediction maps were compared against the ground-truth labels within the ROIs. 
From this comparison, we computed statistical performance metrics such as F1-score, kappa coefficient, and overall accuracy, summarized in Table~III.

Additionally, to assess performance across all decision thresholds, we computed ROC curves and the area under the curve (AUC) for each method in Fig. 9.

@@

\begin{quote}
\textcolor{blue}{These additions appear in Section~IV-C of the revised manuscript.
}
\end{quote}
\end{responsebox}

\vspace{9em}

\begin{reviewbox}{Comment 6}
6. The color bars and the font size of the texts needs to be improved so that it can be clearly visible.
\end{reviewbox}

\begin{responsebox}{Response}
We thank the reviewer for this suggestion. In response, we improved Figures~6–8 (now Figs.~5–7) by increasing the font size of all labels and standardizing the size and placement of the color bars to enhance clarity and readability.

Furthermore, we unified the colormap used in the $p$-value maps across all three scenes to ensure visual consistency and facilitate direct comparison between results.

Additionally, we added the corresponding optical images alongside each SAR image to better support the visual interpretation of heterogeneous regions.

### 2025-06-03 Text/Articles/RENYI_ENTROPY_SAR_2025/R1-Renyi-entropy-sar.qmd
Final edits to R1 and response letters
Unify p-value map colormaps

<!-- ACF Why three different colour tables? -->
```{=latex}
\begin{figure*}[hbt]
    \centering
    \begin{subfigure}{0.145\textwidth}
        \includegraphics[width=\linewidth]{./Figures-R1/london_roi.png}
        \caption{SAR image}
        \label{fig:london-sar}
    \end{subfigure}
    \hspace{0.00001\textwidth}
    \begin{subfigure}{0.142\textwidth}
        \includegraphics[width=\linewidth]{./Figures-R1/london_optical.png}
        \caption{Optical image}
        \label{fig:london-optical}
    \end{subfigure}
    \hspace{0.00001\textwidth}
    \begin{subfigure}{0.179\textwidth}
        \includegraphics[width=\linewidth]{./Figures-R1/p-values_renyi-london_E.png}
        \caption{$p$-value map for $\small{S_{\widetilde{H}_{\lambda}}}$}
        \label{fig:london-renyi}
    \end{subfigure}
    \hspace{0.00001\textwidth}
    \begin{subfigure}{0.145\textwidth}
        \includegraphics[width=\linewidth]{./Figures-R1/H_005_london_renyi_L1_.png}
        \caption{Binary map for $\small{S_{\widetilde{H}_{\lambda}}}$}
        \label{fig:london-005-renyi}
    \end{subfigure}
   \hspace{0.00001\textwidth}
    \begin{subfigure}{0.18\textwidth}
        \includegraphics[width=\linewidth]{./Figures-R1/p-values_shannon-london_E.png}
        \caption{$p$-value map for $\tiny{S_{\widetilde{H}_{\text{AO}}}}$ }
        \label{fig:london-shann}
    \end{subfigure}
    \hspace{0.00001\textwidth}
    \begin{subfigure}{0.145\textwidth}
        \includegraphics[width=\linewidth]{./Figures-R1/H_005_london_shannon.png}
        \caption{Binary map for $\small{S_{\widetilde{H}_{\text{AO}}}}$}
        \label{fig:london-005-shann}
    \end{subfigure}
    \caption{Detection of heterogeneous areas in a SAR image over London, UK: comparison with the tests $\small{S_{\widetilde{H}_{\lambda}}}$ and $\small{S_{\widetilde{H}_{\text{AO}}}}$.}
    \label{fig:london}
\end{figure*}
```

@@

>>>>

<!-- ACF Why three different colour tables? -->

<!-- New color  Palettes viridis B (inferno) -->
```{=latex}
\begin{figure*}[hbt]
    \centering
    \begin{subfigure}{0.145\textwidth}
        \includegraphics[width=\linewidth]{./Figures-R1/london_roi.png}
        \caption{SAR image}
        \label{fig:london-sar}
    \end{subfigure}
    \hspace{0.00001\textwidth}
    \begin{subfigure}{0.142\textwidth}
        \includegraphics[width=\linewidth]{./Figures-R1/london_optical.png}
        \caption{Optical image}
        \label{fig:london-optical}
    \end{subfigure}
    \hspace{0.00001\textwidth}
    \begin{subfigure}{0.179\textwidth}
        \includegraphics[width=\linewidth]{./Figures-R1/p-values_renyi-london_B1.png}
        \caption{$p$-value map for $\small{S_{\widetilde{H}_{\lambda}}}$}
        \label{fig:london-renyi}
    \end{subfigure}
    \hspace{0.00001\textwidth}
    \begin{subfigure}{0.145\textwidth}
        \includegraphics[width=\linewidth]{./Figures-R1/H_005_london_renyi_L1_.png}
        \caption{Binary map for $\small{S_{\widetilde{H}_{\lambda}}}$}
        \label{fig:london-005-renyi}
    \end{subfigure}
   \hspace{0.00001\textwidth}
    \begin{subfigure}{0.18\textwidth}
        \includegraphics[width=\linewidth]{./Figures-R1/p-values_shannon-london_B1.png}
        \caption{$p$-value map for $\tiny{S_{\widetilde{H}_{\text{AO}}}}$ }
        \label{fig:london-shann}
    \end{subfigure}
    \hspace{0.00001\textwidth}
    \begin{subfigure}{0.145\textwidth}
        \includegraphics[width=\linewidth]{./Figures-R1/H_005_london_shannon.png}
        \caption{Binary map for $\small{S_{\widetilde{H}_{\text{AO}}}}$}
        \label{fig:london-005-shann}
    \end{subfigure}
    \caption{Detection of heterogeneous areas in a SAR image over London, UK: comparison with the tests $\small{S_{\widetilde{H}_{\lambda}}}$ and $\small{S_{\widetilde{H}_{\text{AO}}}}$.}
    \label{fig:london}
\end{figure*}
```

```{=latex}
\begin{figure*}[hbt]
    \centering
        \begin{subfigure}{0.145\textwidth}
        \includegraphics[width=\linewidth]{./Figures-R1/munich_roi.png}
        \caption{SAR image}
        \label{fig:munich-sar}
    \end{subfigure}
    \hspace{0.00001\textwidth}
  \begin{subfigure}{0.137\textwidth}
        \includegraphics[width=\linewidth]{./Figures-R1/munich_optical.png}
        \caption{Optical image}
        \label{fig:munich-optical}
    \end{subfigure}
    \hspace{0.00001\textwidth}
    \begin{subfigure}{0.178\textwidth}
        \includegraphics[width=\linewidth]{./Figures-R1/p-values_renyi-munich-B.png}
        \caption{$p$-value map for $\small{S_{\widetilde{H}_{\lambda}}}$}
        \label{fig:munich-renyi}
    \end{subfigure}
    \hspace{0.00001\textwidth}
    \begin{subfigure}{0.144\textwidth}
        \includegraphics[width=\linewidth]{./Figures-R1/H_005_renyi_munich.png}
        \caption{Binary map for $\small{S_{\widetilde{H}_{\lambda}}}$}
        \label{fig:munich-005-renyi}
    \end{subfigure}
    \hspace{0.00001\textwidth}
    \begin{subfigure}{0.178\textwidth}
        \includegraphics[width=\linewidth]{./Figures-R1/p-values_shannon-munich-B.png}
        \caption{$p$-value map for $S_{\widetilde{H}_{\text{AO}}}$}
        \label{fig:munich-shann}
    \end{subfigure}
   \hspace{0.00001\textwidth}
    \begin{subfigure}{0.144\textwidth}
        \includegraphics[width=\linewidth]{./Figures-R1/H_005_shannon_munich.png}
        \caption{Binary map for $\small{S_{\widetilde{H}_{\text{AO}}}}$}
        \label{fig:munich-005-shann}
    \end{subfigure}
    \caption{Detection of heterogeneous areas in a SAR image over Munich, Germany: comparison with the tests $\small{S_{\widetilde{H}_{\lambda}}}$ and $\small{S_{\widetilde{H}_{\text{AO}}}}$.}
    \label{fig:munich}
\end{figure*}

<!-- New color  Palettes viridis H (turbo) -->
```{=latex}
\begin{figure*}[hbt]
    \centering
    \begin{subfigure}{0.145\textwidth}
        \includegraphics[width=\linewidth]{./Figures-R1/london_roi.png}
        \caption{SAR image}
        \label{fig:london-sar}
    \end{subfigure}
    \hspace{0.00001\textwidth}
    \begin{subfigure}{0.142\textwidth}
        \includegraphics[width=\linewidth]{./Figures-R1/london_optical.png}
        \caption{Optical image}
        \label{fig:london-optical}
    \end{subfigure}
    \hspace{0.00001\textwidth}
    \begin{subfigure}{0.179\textwidth}
        \includegraphics[width=\linewidth]{./Figures-R1/p-values_renyi-london_H1.png}
        \caption{$p$-value map for $\small{S_{\widetilde{H}_{\lambda}}}$}
        \label{fig:london-renyi}
    \end{subfigure}
    \hspace{0.00001\textwidth}
    \begin{subfigure}{0.145\textwidth}
        \includegraphics[width=\linewidth]{./Figures-R1/H_005_london_renyi_L1_.png}
        \caption{Binary map for $\small{S_{\widetilde{H}_{\lambda}}}$}
        \label{fig:london-005-renyi}
    \end{subfigure}
   \hspace{0.00001\textwidth}
    \begin{subfigure}{0.18\textwidth}
        \includegraphics[width=\linewidth]{./Figures-R1/p-values_shannon-london_H1.png}
        \caption{$p$-value map for $\tiny{S_{\widetilde{H}_{\text{AO}}}}$ }
        \label{fig:london-shann}
    \end{subfigure}
    \hspace{0.00001\textwidth}
    \begin{subfigure}{0.145\textwidth}
        \includegraphics[width=\linewidth]{./Figures-R1/H_005_london_shannon.png}
        \caption{Binary map for $\small{S_{\widetilde{H}_{\text{AO}}}}$}
        \label{fig:london-005-shann}
    \end{subfigure}
    \caption{Detection of heterogeneous areas in a SAR image over London, UK: comparison with the tests $\small{S_{\widetilde{H}_{\lambda}}}$ and $\small{S_{\widetilde{H}_{\text{AO}}}}$.}
    \label{fig:london}
\end{figure*}
```

@@

```{=latex}
\begin{figure*}[hbt]
    \centering
        \begin{subfigure}{0.145\textwidth}
        \includegraphics[width=\linewidth]{./Figures-R1/dublin_roi3.png}
        \caption{SAR image}
        \label{fig:dublin-sar}
    \end{subfigure}
    \hspace{0.00001\textwidth}
\begin{subfigure}{0.145\textwidth}
        \includegraphics[width=\linewidth]{./Figures-R1/dublin.png}
        \caption{Optical image}
        \label{fig:dublin-optical}
    \end{subfigure}
    \hspace{0.00001\textwidth}
    \begin{subfigure}{0.178\textwidth}
        \includegraphics[width=\linewidth]{./Figures-R1/p-values_renyi-dublin-H1.png}
        \caption{$p$-value map $\small{S_{\widetilde{H}_{\lambda}}}$}
        \label{fig:dublin-renyi}
    \end{subfigure}
    \hspace{0.00001\textwidth}
    \begin{subfigure}{0.144\textwidth}
        \includegraphics[width=\linewidth]{./Figures-R1/H_005_dublin_renyi.png}
        \caption{Binary map $\small{S_{\widetilde{H}_{\lambda}}}$}
        \label{fig:dublin-005-renyi}
    \end{subfigure}
    \hspace{0.00001\textwidth}
    \begin{subfigure}{0.178\textwidth}
        \includegraphics[width=\linewidth]{./Figures-R1/p-values_shannon-dublin-H1.png}
        \caption{$p$-value $\small{S_{\widetilde{H}_{\text{AO}}}}$}
        \label{fig:dublin-shann}
    \end{subfigure}
    \hspace{0.0000001\textwidth}
    \begin{subfigure}{0.144\textwidth}
        \includegraphics[width=\linewidth]{./Figures-R1/H_005_dublin-shannon.png}
        \caption{Binary $\small{S_{\widetilde{H}_{\text{AO}}}}$}
        \label{fig:dublin-005-shann}
    \end{subfigure}
    \caption{Detection of heterogeneous areas in a SAR image over Dublin Port, Ireland: comparison with the tests $\small{S_{\widetilde{H}_{\lambda}}}$ and $\small{S_{\widetilde{H}_{\text{AO}}}}$.}
    \label{fig:dublin}
\end{figure*}
